Approximating the matrix exponential derivative of the CTMC generator matrix, with application to global SARS-CoV-2

### 2023-03-16 manuscript/main.tex
need tree then done

We generate 8 million MCMC samples in this manner, saving 1 in 10,000 Markov chain states, in order to guarantee a minimum effective sample size (ESS) greater than 100.  By far, the worst-mixing parameter  is the global scale $\tau$, which obtains an ESS of 185.  The three fixed-effects regression coefficients $\theta_1$, $\theta_2$ and $\theta_3$ obtain ESS's of 721, 721 and 644, respectively.  The median and minimum ESS for the 1,892 random-effects $\epsilon_{ss'}$ are 721 and 190, respectively.  We note that, after thinning and removing burn-in, the sample only consists of 721 states, so ESS of 721 implies negligible autocorrelation between samples.  Figure \ref{fig:post} presents

\begin{itemize}
	\item Figure 2
	\item Tree
\end{itemize}

>>>>

We generate 8 million MCMC samples in this manner, saving 1 in 10,000 Markov chain states, in order to guarantee a minimum effective sample size (ESS) greater than 100.  By far, the worst-mixing parameter  is the global scale $\tau$, which obtains an ESS of 185.  The three fixed-effects regression coefficients $\theta_1$, $\theta_2$ and $\theta_3$ obtain ESS's of 721, 721 and 644, respectively.  The median and minimum ESS for the 1,892 random-effects $\epsilon_{ss'}$ are 721 and 190, respectively.  We note that, after thinning and removing burn-in, the sample only consists of 721 states, so ESS of 721 implies negligible autocorrelation between samples.  The left plot of Figure \ref{fig:post} presents posterior densities for the fixed-effect coefficients from regression \eqref{eq:mixed}.  The air traffic proximity, intracontinental proximity and Hubei asymmetry coefficients have posterior means and 95\% credible intervals of 0.76 (0.50, 1.03), 0.04 (-0.04, 0.12) and 0.27 (-0.39, 0.78), respectively. While these results suggest a positive association between each predictor matrix and the infinitesimal generator matrix $\QQ$, the only predictor with a statistically significant association is air traffic proximity.  This result agrees with a previous phylogeographic analysis of the global spread of influenza \citep{holbrook2021massive}.  The right plot of Figure \ref{fig:post} presents the posterior mean for generator $\QQ$.  As one may expect, the matrix looks similar to that of the air traffic predictor matrix in Figure \ref{fig:FE}, but one may also see the influence of the Hubei asymmetry in, e.g., the squares corresponding to travel between Guangdong and Hubei provinces.

\begin{itemize}
	\item tree
\end{itemize}
